\documentclass[11pt,a4paper]{article}
\usepackage[margin=2.5cm]{geometry}
\usepackage{parskip}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{microtype}
\usepackage{hyperref}
\hypersetup{colorlinks=true, urlcolor=blue, linkcolor=black, citecolor=black}

% Pull macros from the same source as the thesis so the headline numbers
% never drift from what the analysis pipeline actually produced.
\input{generated/thesis_values}
\input{generated/icf_values}

\pagestyle{empty}

\begin{document}

\hfill 4 June 2026

\bigskip

{\LARGE\bfseries Young women taking SSRI antidepressants may be at increased risk of suicidal behavior}

\medskip

%{\large Population-wide Swedish registry study of \studyN{} young people identifies
%sex as the dominant axis of variation in early-period harm.}

%\bigskip

\noindent\textbf{Stockholm, 4 June 2026.} Initiating a selective serotonin
reuptake inhibitor (SSRI) within 28~days of a first depression diagnosis was
estimated to raise 12-week suicidal-behavior risk by \ittRD~percentage
points (95\% CI \ittRDLower, \ittRDUpper) among Swedish young people aged
6 to 24, with the increase roughly \icfCATEratio$\times$ as large in
females as in males, according to a new master's thesis from Karolinska
Institutet.

The analysis followed \studyN{} individuals with a first depression
diagnosis between 2006 and 2019, drawing on Sweden's national health,
prescription, and population registers. It used target-trial emulation -- a
method that derives trial-like causal estimates from observational data --
together with iterative causal forests, a recently developed
machine-learning technique that searches for patient subgroups who respond
differently to treatment.

After adjusting for differences in patient characteristics between those who
did and did not start treatment, the algorithm discovered a meaningful
distinction between male and female patients. The estimated 12-week absolute
risk increase was \icfSGACATE~percentage points in males (95\% CI
\icfSGACILower, \icfSGACIUpper) and \icfSGBCATE~percentage points in
females (\icfSGBCILower, \icfSGBCIUpper). Searches across additional
diagnostic codes and prescription patterns produced no further reproducible
subgroups.

``The risk increase in the first 12~weeks is small overall, but it is
significantly larger in female patients than in male patients,'' said Ville
Sebastian Olsson, the thesis author. ``Sex should be taken into consideration
when starting one of these drugs, not just age.''

The FDA first issued a black-box warning on SSRIs in 2004 after a
meta-analysis of pediatric trials found roughly a two-fold higher rate of
suicidal ideation or attempts in treated children and adolescents. That
warning stratifies on age, not on sex. The new thesis adds sex as a
candidate additional axis of caution in a national real-world cohort more
than two decades after the original trial-based finding.

The thesis emphasizes that the contrast is between starting an SSRI and not
starting one, not between SSRIs and alternative treatments such as
psychotherapy. It is also noted that the absolute increase is small (roughly one
additional event per 100 patients treated over 12~weeks). Replication
in independent registries, longer follow-up, and an active-comparator
analysis are listed as the natural next steps.

\medskip

\noindent\textbf{Contact:}\\
Ville Sebastian Olsson, thesis author\\
Department of Medical Epidemiology and Biostatistics, Karolinska Institutet\\
Email: \texttt{ville.sebastian.olsson@ki.se}\\

\section*{Statement on the use of generative AI}
I used OpenAI ChatGPT to assist in providing feedback on the written text and
helping me refine it.

\end{document}
